---
title: "General Social Survey"
author: "Daniel"
date: "Fall 2023"
output: 
  html_document:
    number_sections: true
    toc: true
---


```{r include=FALSE,echo=FALSE}
require(tigerstats)
require(tidyverse)
require(foreign)
source("GrabData.R")
```


#  Question 1

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

The data in the dataframe GSSdata is from the 2018 General Social Survey. The first blocks of R-code has selected down a subset of the data to just 16 variables. It has further removed unwanted factor levels in much of the data. Examine the code in the GrabData.R file to see what it is doing. Some of the variables are categorical and others are numerical. Be sure to do a variable analysis before tackling each question.  
First question - Is opinion on the death penalty (CAPPUN) independent of gun ownership (OWNGUN)?

$H_0$ There is no relationship between gun ownership and opinion on the death penalty 

$H_A$ There is a relationship between gun ownership and opinion on the death penalty 

## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->


Both variables are categorical so we use Cat~Cat helper file

##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
barchartGC(~OWNGUN + CAPPUN,data=GSSdata)
barchartGC(~OWNGUN + CAPPUN,data=GSSdata, type="percent")
```

By the graph, those who own a gun are more likely to favor the death penalty  

#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
table1 <- xtabs(~OWNGUN + CAPPUN,data=GSSdata)
rowPerc(table1)
colPerc(table1)
```

This summary that among those with guns are more likely to be in favor of capital punishment, and those who do not own guns are less likely to be in favor of capital punishment.

### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisq.test(table1)
chisqtestGC(table1)
```

The chisq test, we are given a p-value of 0.02022 which is less than .05, which means we reject the null hypothesis and there is a relationship between owning a gun and favor of capital punishment.

```{r}
fisher.test(table1)
```

With a fisher test, we get a p-value of 0.01651, which is less than .05, which means we reject the null hypothesis. There is a relationship between owning a gun and favor of capital punishment. People who own a gun are 2.3 times more likely to favor capital punishment that those who do not own a gun.

#  Question 2

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

Is there a difference in years of education (EDUC) between the those who did or did not vote in 2012 (VOTE12)?

$H_0$ There is no difference in years of education and weather those did/didn't vote in 2012

$H_A$ There is a difference in years of education and weather those did/didn't vote in 2012

## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

We have a numerical variable and a categorical variable, therefore we use Numerical~Cat helper file   

##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
ggplot(GSSdata, aes(x = VOTE12, y = EDUC)) +
  geom_boxplot() + 
  labs(
    title = "VOTE12 vs EDUC",
    x = "VOTE12",
    y = "EDUC"
  )

```

The graph shows that those with more education, are more likely to vote as opposed to those with less education.

#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
#Calculate summary stats
summary_stats <-  GSSdata %>%
  group_by(VOTE12) %>%
  summarise(
    Mean = mean(EDUC),
    Median = median(EDUC),
    SD = sd(EDUC)
  )
summary_stats
```

The mean of those who voted was slightly higher than those that did not vote, close to 3 point difference, and the median is the actual 3 point difference. This shows that there is a coloration between years of education and weather someone votes.

### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
anova_result <- aov(EDUC ~ VOTE12, data = GSSdata)
summary(anova_result)

```

We have a p-value of 4.53e^-7 which is lower than 0.05, which means that we reject the null hypothesis, meaning there is a relationship between years of education and weather people voted or not in 2012

#  Question 3

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

Determine if a persons performance on the wordsum test (WORDSUM) is independent of their level of schooling (DEGREE).

## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

$H_0$ There is no relationship in wordsum test scores and level of schooling

$H_A$ There is a relationship in wordsum test scores and level of schooling

##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->



### Descriptive Results

#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->


# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

